Virtual Element Methods Without Extrinsic Stabilization

### 2023-09-21 stabfreevem.tex
add refs on stab-free VEMs

% The idea in \cite{BerroneBorioMarcon2021} is difficult to be extended to construct stabilization-free virtual element methods in higher dimensions, and
% the analysis is rather elaborate. 
% This motivates us to construct a nonconforming VEM in arbitrary dimension and a conforming VEM in two dimensions in arbitrary degree of polynomial  in a unified way.
We intend to construct VEMs without extrinsic stabilization in arbitrary degree of polynomial, including a nonconforming VEM and a conforming VEM in arbitrary dimension,  in a unified way.
The key to construct VEMs without extrinsic stabilization is to find a finite-dimensional space $\mathbb{V}(K)$ for polytope $K$ and a projector $Q_K$ onto the space $\mathbb{V}(K)$ such that
\begin{enumerate}[(C1)]
\item It holds the norm equivalence 
\begin{equation}\label{intro:gradVknormequiv} 
\|Q_{K}\nabla v\|_{0,K}\eqsim \|\nabla v\|_{0,K} \quad \forall~v\in V_k(K)
\end{equation}
on shape function space $V_k(K)$ of virtual elements;
\item The projection $Q_{K}\nabla v$ is computable based on the degrees of freedom (DoFs) of virtual elements for $v\in V_k(K)$.
\end{enumerate}
The hidden constants in \eqref{intro:gradVknormequiv} are independent of the size of $K$, but depend on the degree of polynomials, and the chunkiness parameter and the geometric dimension of $K$; see Section~\ref{sec:meshcondition} for details.
We can choose $Q_{K}$ as the $L^2$-orthogonal projector with respect to the
inner product $(\cdot, \cdot)_K$. The norm equivalence
\eqref{intro:gradVknormequiv} implies that the space $\mathbb{V}(K)$ should be
sufficiently large compared with the virtual element space $V_k(K)$.  In
standard virtual element methods, $Q_{k-1}^{K}\nabla v$
\cite{BeiraodaVeigaBrezziMariniRusso2016} or $\nabla\Pi_k^{K}v$
\cite{BeiraoBrezziCangianiManziniEtAl2013,BeiraoBrezziMariniRusso2014,AhmadAlsaediBrezziMariniEtAl2013,AyusodeDiosLipnikovManzini2016}
are used, where $Q_{k-1}^{K}$ is the $L^2$-orthogonal projector onto the
$(k-1)$-th order polynomial space $\mathbb P_{k-1}(K; \mathbb{R}^d)$, and
$\Pi_k^{K}$ is the $H^1$ projection operator onto the $k$-th order polynomial
space $\mathbb P_{k}(K)$.  While only
\[
\|Q_{k-1}^{K}\nabla v\|_{0,K}\lesssim \|\nabla v\|_{0,K}, \quad \|\nabla\Pi_k^{K}v\|_{0,K}\lesssim \|\nabla v\|_{0,K}
\]
hold
rather than the norm equivalence \eqref{intro:gradVknormequiv}, then an additional stabilization term is usually required to ensure the coercivity of the discrete bilinear form.
To remove the additional stabilization term, based on a regular simplicial tessellation of polytope $K$, we employ $k$-th order or $(k-1)$-th order $H(\div)$-conforming macro finite elements as $\mathbb{V}(K)$ in this paper, and keep the virtual element space $V_k(K)$ as the usual ones.
% $\mathbb{V}(K)$ is taken as $\mathbb P_{l}(K; \mathbb{R}^d)$ with large enough $l$ even for the lowest order case $k=1$ in \cite{BerroneBorioMarcon2021,BerroneBorioMarcon2022,DAltriMirandaPatrunoSacco2021}, and the virtual element space $V_k(K)$ has to be modified accordingly to make $Q_{l}^{K}\nabla v$ computable.
% Instead, based on a regular simplicial tessellation of polytope $K$, we employ $k$-th order or $(k-1)$-th order $H(\div)$-conforming macro finite elements as $\mathbb{V}(K)$ in this paper, and keep the virtual element space $V_k(K)$ as the usual ones.

>>>>

A linear VEM without extrinsic stabilization, based on a higher order polynomial projection of the gradient of virtual element functions, is devised for the Poisson equation in two dimensions in \cite{BerroneBorioMarcon2021}, where the degree of polynomial used in the projection depends on the number of vertices of the polygon and generally the geometry of the polygon. 
Numerical examples in \cite{BerroneBorioMarcon2022} show that the VEM in \cite{BerroneBorioMarcon2021} outperforms the standard one in \cite{BeiraodaVeigaBrezziMariniRusso2016} for anisotropic elliptic problems on general convex polygonal meshes.
% The stabilization-free VEM in \cite{BerroneBorioMarcon2021} is then applied to solve the Laplacian eigenvalue problem \cite{MengWangBuMei2022}. 
% Along this line, similar stabilization-free virtual element methods are devised for a plane elasticity problem in~\cite{DAltriMirandaPatrunoSacco2021,ChenSukumar2022,ChenSukumar2023}. 
The idea in \cite{BerroneBorioMarcon2021} is difficult to be extended to construct VEMs without extrinsic stabilization in higher dimensions, and
the analysis is rather elaborate. 
This motivates us to construct VEMs without extrinsic stabilization in arbitrary dimension and arbitrary degree of polynomial in a unified way.

% The idea in \cite{BerroneBorioMarcon2021} is difficult to be extended to construct stabilization-free virtual element methods in higher dimensions, and
% the analysis is rather elaborate. 
% This motivates us to construct a nonconforming VEM in arbitrary dimension and a conforming VEM in two dimensions in arbitrary degree of polynomial  in a unified way.
% We intend to construct VEMs without extrinsic stabilization in arbitrary degree of polynomial, including a nonconforming VEM and a conforming VEM in arbitrary dimension,  in a unified way.

The key to construct VEMs without extrinsic stabilization is to find a finite-dimensional space $\mathbb{V}(K)$ for polytope $K$ and a projector $Q_K$ onto the space $\mathbb{V}(K)$ such that
\begin{enumerate}[(C1)]
\item It holds the norm equivalence 
\begin{equation}\label{intro:gradVknormequiv} 
\|Q_{K}\nabla v\|_{0,K}\eqsim \|\nabla v\|_{0,K} \quad \forall~v\in V_k(K)
\end{equation}
on shape function space $V_k(K)$ of virtual elements;
\item The projection $Q_{K}\nabla v$ is computable based on the degrees of freedom (DoFs) of virtual elements for $v\in V_k(K)$.
\end{enumerate}
The hidden constants in \eqref{intro:gradVknormequiv} are independent of the size of $K$, but depend on the degree of polynomials, and the chunkiness parameter and the geometric dimension of $K$; see Section~\ref{sec:meshcondition} for details.
We can choose $Q_{K}$ as the $L^2$-orthogonal projector with respect to the
inner product $(\cdot, \cdot)_K$. The norm equivalence
\eqref{intro:gradVknormequiv} implies that the space $\mathbb{V}(K)$ should be
sufficiently large compared with the virtual element space $V_k(K)$.  In
standard virtual element methods, $Q_{k-1}^{K}\nabla v$
\cite{BeiraodaVeigaBrezziMariniRusso2016} or $\nabla\Pi_k^{K}v$
\cite{BeiraoBrezziCangianiManziniEtAl2013,BeiraoBrezziMariniRusso2014,AhmadAlsaediBrezziMariniEtAl2013,AyusodeDiosLipnikovManzini2016}
are used, where $Q_{k-1}^{K}$ is the $L^2$-orthogonal projector onto the
$(k-1)$-th order polynomial space $\mathbb P_{k-1}(K; \mathbb{R}^d)$, and
$\Pi_k^{K}$ is the $H^1$ projection operator onto the $k$-th order polynomial
space $\mathbb P_{k}(K)$.  While only
\[
\|Q_{k-1}^{K}\nabla v\|_{0,K}\lesssim \|\nabla v\|_{0,K}, \quad \|\nabla\Pi_k^{K}v\|_{0,K}\lesssim \|\nabla v\|_{0,K}
\]
hold
rather than the norm equivalence \eqref{intro:gradVknormequiv}, then an additional stabilization term is usually required to ensure the coercivity of the discrete bilinear form.
To remove the additional stabilization term, based on a regular simplicial tessellation of polytope $K$, we employ $k$-th order or $(k-1)$-th order $H(\div)$-conforming macro finite elements as $\mathbb{V}(K)$ in this paper, and keep the virtual element space $V_k(K)$ as the usual ones.
% $\mathbb{V}(K)$ is taken as $\mathbb P_{l}(K; \mathbb{R}^d)$ with large enough $l$ even for the lowest order case $k=1$ in \cite{BerroneBorioMarcon2021,BerroneBorioMarcon2022,DAltriMirandaPatrunoSacco2021}, and the virtual element space $V_k(K)$ has to be modified accordingly to make $Q_{l}^{K}\nabla v$ computable.
% Instead, based on a regular simplicial tessellation of polytope $K$, we employ $k$-th order or $(k-1)$-th order $H(\div)$-conforming macro finite elements as $\mathbb{V}(K)$ in this paper, and keep the virtual element space $V_k(K)$ as the usual ones.
